Multi-fluids for Representing Sub-filter-scale Convection

### 2020-07-03 submit3/manuscript2.tex
First draft of revisions and responses to second set of reviews from reviewer 1. Still to respond to reviewer 2. I have currently not followed all of Dan's suggestions because Dan's work should go into Dan's first author papers. I still need to reponsd to and revise based on reviewer 2. I also need to add a couple of lines about the boundary layer depth for the RCE test case.

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
%\usepackage{soul}
\soulregister\ref7
\soulregister\cite7
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%
\addeditor{HW}
\addeditor{WM}

@@

\authors
{
    Hilary Weller\affil{1}\thanks{Meteorology, University of Reading, UK}
    ,
    William McIntyre\affil{1}
    ,
    Daniel Shipley\affil{1}
    , and 
    Peter Clark\affil{1}
}

@@

\begin{keypoints}
\item Multi-fluid modelling enables net mass transport by parameterised convection and non-equlibrium.
\item Multi-fluid modelling means a new dynamical core rather than a parameterisation for an existing core
\item Closure for the pressure difference between fluids and entrainment and detrainment are provided
\end{keypoints}

\begin{abstract}
Multi-fluid modelling is a promising new method of representing sub
and near grid-scale convection allowing for consistent and accurate
treatment of net mass transport by convection \add[HW]{and of} non-equilibrium \add[HW]{dynamics} \remove[HW]{and turbulence}. The air is partitioned into \change[HW]{convecting and non-convecting}{two or more}
fluids \add[HW]{which may represent, for example, updrafts and the stable environment.} \change[HW]{each with their}{Each fluid has its} own velocity, temperature and constituents
with separate equations of motion. \remove[HW]{can be solved for each fluid,
thus representing features that cannot be resolved in a single-fluid
at coarse resolution.}

This paper presents \change[HW]{multi}{two}-fluid Boussinesq equations for representing under-resolved dry convection
\add[HW]{with sinking and $w=0$ air in fluid 0 and rising air in fluid 1}. \change[HW]{We present a new}{A} model for entrainment
and detrainment based on divergence \remove[HW]{which} leads to excellent representation of the convective area fraction.
Previous multi-fluid modelling of convection has
\change[HW]{assumed that all fluids share the same pressure.}
{used the same pressure for both fluids.}
\change[HW]{We show that this is}{This is shown to be} a bad \change[HW]{assumption}{approximation} and \remove[HW]{we propose and validate} a model
for the pressure difference between the fluids based on divergence \add[HW]{is presented}.

Two vertical slice test cases are \change[HW]{used}{developed} to \add[HW]{tune parameters and to} \change[HW]{validate}{evaluate} the \change[HW]{multi}{two}-fluid \change[HW]{model}{equations for representing dry, two-dimensional, sub-grid-scale convection}; a buoyant rising bubble and radiative convective equilibrium. \add[HW]{These}
are \add[HW]{first} simulated at high resolution and conditionally averaged based
on the sign of the vertical velocity. \add[HW]{The test cases are next simulated with the two-fluid model which} \remove[HW]{The multi-fluid model} reproduces
the mean properties of the rising and falling fluids in \change[HW]{a single}{one} column. 
\end{abstract}

\section*{Plain Language Summary}
Clouds and buoyant convection are often smaller than the grid size in weather and climate prediction models but they are of central importance. Therefore their effects are parameterised -- their interactions with the larger scales are estimated based on properties of the larger scales. Convection parameterisations are a large source of error in models of the atmosphere despite decades of effort to improve them. An important assumption in most convection parameterisations is that convection moves heat and moisture upwards but not mass. This makes it much simpler to incorporate convection parameterisations into existing models. Multi-fluid modelling will enable mass transport by convection but the approach means a whole new atmospheric model rather than a stand alone routine that can be incorporated into an existing model. This paper provides some closures necessary for multi-fluid modelling and \change[HW]{validation}{evaluation} of the technique for two dimensional dry convection.

@@

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\add[HW]{The multi-fluids approach has an advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments directly (mean properties of eg. updraft and environment) rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for sub-fluid means could be less sensitive to closure assumptions.}
\remove{with the same numerics and physics to prognose properties of separate parts of the fluid, for example updrafts, the stable environment and downdrafts dynamics}. \citeA{Yano14} derived similar
equations \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme \remove[HW]{also} uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}
{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}
\change[HW]{Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids}
{The disadvantage of the multi-fluids approach is that it} requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of \change[HW]{the multi}{two}-fluid equations
\change[HW]{assuming two fluids that share}{using} the same pressure
\add[HW]{for both fluids}.
\remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations \add[HW]{because} convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore\change[WM]{ for simplicity we focus on the}{, the}
\remove[HW]{multi-fluid} Boussinesq equations without background stratification \add[WM]{are chosen for simplicity}:

\note[HW]{Additional material until the start of 2.1}
\begin{eqnarray}
\frac{\partial \mathbf{u}}{\partial t} + \nabla\cdot(\mathbf{u}\mathbf{u})
+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{\partial b}{\partial t} + \nabla\cdot(\mathbf{u}b)
 & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}.

\begin{table}
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$
\tabularnewline
\add[HW]{$\mathbf{u}$} & 
\add[HW]{Velocity of the single-fluid equations  ($\text{m}\text{s}^{-1}$) and velocity averaged over all fluids of the multi-fluid equations}
\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)
\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)
\tabularnewline
\add[HW]{$P$} & \add[HW]{Non-hydrostatic pressure of the single-fluid equations $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$) and pressure averaged over all fluids of the multi-fluid equations}
\tabularnewline
$P_i$ & Non-hydrostatic pressure of fluid $i$
\tabularnewline
\add[HW]{$p_i$} & \add[HW]{$P_i - P$}
\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$
\tabularnewline
$b$ & Buoyancy of single-fluid equations defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$) \add[HW]{and buoyancy averaged over all fluids of the multi-fluid equations}
\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$
\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$\gamma$ & \change[HW]{Coefficient for setting the pressure local for each fluid}{Compressibility for calculating $p_i$} ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1
\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
\add[WM]{$\mathbf{k}$} & \add[WM]{Unit vector in the $z$-direction.}
\tabularnewline
\end{tabular}
\caption{Definitions of variables for the single- and multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and 0 elsewhere. 
For the derivation of the conditionally averaged equations \change[WM]{we will not consider relabelling of fluid}{the relabelling of fluids will not be considered,} \add[HW]{and so} $I_i$ satisfies:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
\change[WM]{Applying conditional volume average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$}{By applying a conditional volume average (denoted by the overbar), the fluid $i$ volume fractions, velocities and buoyancies are defined by}:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
\change[WM]{As we are considering the Boussinesq equations we assume}{As the system is governed by the Boussinesq equations, it is assumed} that variations in density only affect the \change[WM]{buoyancy terms}{definition of the buoyancy terms (table \protect\ref{tab:defns})} so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

To derive a transport equation for $\sigma_i$\change[WM]{ we combine (\protect\ref{eq:LagrangianLabel})}{, (\protect\ref{eq:LagrangianLabel}) is combined} with $I_i$ times (\ref{eq:singleDivFree}) and volume average. After permuting operators with the averaging \cite<as described by>{TWV+18}\add[WM]{,} and using the definitions of $\sigma_i$ and $\mathbf{u}_i$\add[WM]{,} \remove[WM]{we get} the evolution equation for $\sigma_i$\add[WM]{ is obtained}:
\begin{equation}
\frac{\partial\sigma_{i}}{\partial t}+\nabla\cdot\sigma_{i}\mathbf{u}_{i}  =  \sum_{j\ne i}\left\{ \sigma_{j}S_{ji}-\sigma_{i}S_{ij}\right\}
\label{eq:sigma}
\end{equation}
which also includes transfers between fluids with $\sigma_{i}S_{ij}$ being the transfer from fluid $i$ to $j$.

Next \change[WM]{we will consider conditionally averaging the terms of }{the conditional averaging of terms in} the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations\add[WM]{ is considered}. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.
%
\note[HW]{Dan, can we do better than this?}
\change[WM]{Lots of}{Many} terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. \change[WM]{Therefore we will use simply}{The conditionally averaged diffusion terms are therefore assumed to behave like}:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
\change[WM]{Unlike Thuburn et al. (2018) we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:}{The pressures for each fluid, $P_i$, are defined to satisfy:}
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i
\end{equation}
\add[WM]{which differs from \protect\citeA{TWV+18} who use a shared fluid pressure.}
%
Combining \change[WM]{these}{the above terms} gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

>>>>

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
%\usepackage[inline]{trackchanges} %for better track changes. finalnew option
\usepackage[finalnew]{trackchanges}
\soulregister\ref7
\soulregister\cite7
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%
\addeditor{HW}
\addeditor{WM}

@@

\authors
{
    Hilary Weller\affil{1}\thanks{Meteorology, University of Reading, UK}
    ,
    William McIntyre\affil{1}
    ,
    Daniel Shipley\affil{1}
}

@@

\begin{keypoints}
\item Multi-fluid modelling enables net mass transport by \change[HW]{parameterised}{sub-grid-scale} convection and non-equlibrium.
\item Multi-fluid modelling means a new dynamical core rather than a parameterisation for an existing core
\item Closures for the pressure difference between fluids and entrainment and detrainment are provided
\end{keypoints}

\begin{abstract}
Multi-fluid modelling is a promising new method of representing sub- and near-grid-scale convection allowing for consistent and accurate
treatment of net mass transport by convection \add[HW]{and of} non-equilibrium \add[HW]{dynamics} \remove[HW]{and turbulence}. The air is partitioned into \change[HW]{convecting and non-convecting}{two or more}
fluids \add[HW]{which may represent, for example, updrafts and the stable environment.} \change[HW]{each with their}{Each fluid has its} own velocity, temperature and constituents
with separate equations of motion. \remove[HW]{can be solved for each fluid,
thus representing features that cannot be resolved in a single-fluid
at coarse resolution.}

This paper presents \change[HW]{multi}{two}-fluid Boussinesq equations for representing under-resolved dry convection
\add[HW]{with sinking and $w=0$ air in fluid 0 and rising air in fluid 1}. \change[HW]{We present a new}{A} model for entrainment
and detrainment based on divergence \remove[HW]{which} leads to excellent representation of the convective area fraction.
Previous multi-fluid modelling of convection has
\change[HW]{assumed that all fluids share the same pressure.}{used the same pressure for both fluids.}
\change[HW]{We show that this is}{This is shown to be} a bad \change[HW]{assumption}{approximation} and \remove[HW]{we propose and validate} a model
for the pressure difference between the fluids based on divergence \add[HW]{is presented}.

Two vertical slice test cases are \change[HW]{used}{developed} to \add[HW]{tune parameters and to} \change[HW]{validate}{evaluate} the \change[HW]{multi}{two}-fluid \change[HW]{model}{equations for representing dry, sub-grid-scale convection}: a buoyant rising bubble and radiative convective equilibrium. \add[HW]{These}
are \add[HW]{first} simulated at high resolution and conditionally averaged based
on the sign of the vertical velocity. \add[HW]{The test cases are next simulated with the two-fluid model which} \remove[HW]{The multi-fluid model} reproduces
the mean properties of the rising and falling fluids in \change[HW]{a single}{one} column. 
\end{abstract}

\section*{Plain Language Summary}
Clouds and buoyant convection are often smaller than the grid size in weather and climate prediction models but they are of central importance. Therefore their effects are parameterised -- their interactions with the larger scales are estimated based on properties of the larger scales. Convection parameterisations are a large source of error in models of the atmosphere despite decades of effort to improve them. 
\change[HW]{An important assumption in most convection parameterisations is that convection moves heat and moisture upwards but not mass. This makes }{Some unrealistic assumptions remain such as that convection does not transport mass upwards, convection is in equilibrium with the surroundings and that the variations in the horizontal direction are not relevant. These make}
it much simpler to incorporate convection parameterisations into existing models. 
Multi-fluid modelling
\change[HW]{will enable mass transport by convection}{does not rely on these assumptions}
but the approach means a whole new atmospheric model rather than a stand alone routine that can be incorporated into an existing model. This paper provides some closures necessary for multi-fluid modelling and \change[HW]{validation}{evaluation} of the technique for two dimensional dry convection.

@@

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\add[HW]{The multi-fluids approach has a potential advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments directly (mean properties of e.g. updraft and environment) rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for first-order moments in each fluid could be less sensitive to closure assumptions.}
\add[HW]{The multi-fluids approach has an advantage over prognostic plume models such as \protect\citeA{GG05} as multi-fluids enables the mass transported by convection to feed back onto the continuity equation.}
\remove{with the same numerics and physics to prognose properties of separate parts of the fluid, for example updrafts, the stable environment and downdrafts dynamics}. \citeA{Yano14} derived similar
equations to \citeA{TWV+18} \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme \remove[HW]{also} uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}

\change[HW]{Multi-fluids promises big improvements in the representation of sub-grid-scale convection and unification with the representation of turbulence, particularly at the grey zone. However, there is a long way to go, particularly as multi-fluids}{The disadvantage of the multi-fluids approach is that it} requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of \change[HW]{the multi}{two}-fluid equations
\change[HW]{assuming two fluids that share}{using} the same pressure
\add[HW]{for both fluids}.
\remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper, \remove[HW]{we show that there are} significant
and sustained pressure differences between the fluids \add[HW]{are shown to exist} (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations \add[HW]{because} convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Boussinesq equations. Therefore\change[WM]{ for simplicity we focus on the}{, the}
\remove[HW]{multi-fluid} Boussinesq equations without background stratification \add[WM]{are chosen for simplicity}:
\begin{eqnarray}
\frac{\partial \mathbf{u}}{\partial t} + \nabla\cdot(\mathbf{u}\mathbf{u})
+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{\partial b}{\partial t} + \nabla\cdot(\mathbf{u}b)
 & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}. \add[HW]{Note that the ``pressure'' variable, $P$, is actually pressure divided by a constant reference density.}

\begin{table}
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$
\tabularnewline
\add[HW]{$\mathbf{u}$} & 
\add[HW]{Velocity of the single-fluid equations  ($\text{m}\text{s}^{-1}$) and velocity averaged over all fluids of the multi-fluid equations}
\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)
\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Rate of volume fraction transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)
\tabularnewline
\add[HW]{$P$} & \add[HW]{Perturbation pressure of the single-fluid equations $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$) and pressure averaged over all fluids of the multi-fluid equations}
\tabularnewline
$P_i$ & Perturbation pressure of fluid $i$
\tabularnewline
\add[HW]{$p_i$} & \add[HW]{$P_i - P$}
\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$
\tabularnewline
$b$ & Buoyancy of single-fluid equations defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$) \add[HW]{and buoyancy averaged over all fluids of the multi-fluid equations}
\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$
\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$\gamma$ & \change[HW]{Coefficient for setting the pressure local for each fluid}{Compressibility for calculating $p_i$} ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1
\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
\add[WM]{$\mathbf{k}$} & \add[WM]{Unit vector in the $z$-direction.}
\tabularnewline
\end{tabular}
\caption{Definitions of variables for the single- and multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

\note[HW]{Additional material until the start of 2.2}

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels, $I_i$, pick out different fluid types or regions. $I_i=1$ in fluid $i$ and 0 elsewhere. $I_i$ satisfies Lagrangian conservation without transfers:
\begin{equation}
\frac{\partial I_i}{\partial t} + \mathbf{u}\cdot\nabla I_i = 0.
\label{eq:LagrangianLabel}
\end{equation}
It is more straightforward to introduce transfers after volume averaging because the transfer terms for (\ref{eq:LagrangianLabel}) involve delta functions (this will be addressed in future publications). 
By applying a volume average (denoted by the overbar), the fluid $i$ volume fractions, velocities and buoyancies are defined by:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
As the system is governed by the Boussinesq equations, it is assumed that variations in density only affect the definition of the buoyancy terms (table \protect\ref{tab:defns}) so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

To derive a transport equation for $\sigma_i$, (\ref{eq:LagrangianLabel}) is combined with $I_i$ times (\ref{eq:singleDivFree}) and volume averaged. 
It is necessary to assume that
operators permute with the averaging \cite<as described by>{TWV+18}. Transfers between fluids are included at this stage:
\begin{equation}
\frac{\partial\sigma_{i}}{\partial t}+\nabla\cdot\sigma_{i}\mathbf{u}_{i}  =  \sum_{j\ne i}\left\{ \sigma_{j}S_{ji}-\sigma_{i}S_{ij}\right\}
\label{eq:sigma}
\end{equation}
where $\sigma_{i}S_{ij}$ is the transfer rate from fluid $i$ to $j$.

Next the conditional averaging of terms in the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations is considered. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes are represented by constant viscosity diffusion terms for each fluid. The diffusion terms are written so that, in the absence of transfers between fluids and if both fluids are initialised to be identical, the fluids remain identical and $\sigma_i$ acts as a passive tracer:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &\approx& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &\approx& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
The pressures within each fluid are defined as:
\begin{equation}
\sigma_i P_i = \overline{I_i P}.
\end{equation}
The pressure gradient can then be decomposed as:
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i + 
    \left[ P_i\nabla\sigma_i - \overline{P\nabla I_i} \right]
\end{equation}
The terms in square brackets are the difference between the pressure force pushing outward from the fluid calculated using resolved variables and the full resolved version of the same force, both of which are related to drag. They will be represented by a drag, $\mathbf{D}_{ij}$. However this drag is likely to be smaller than the drag that is used, for example, in multi-phase modelling \cite<e.g.>{GBB+07} as it is representing the difference between two representations of drag, rather than the whole drag. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

Combining all terms gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{
     \sigma_{j} S_{ji}\mathbf{u}_{ji}^{T}
   - \sigma_{i} S_{ij}\mathbf{u}_{ij}^{T}
   - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i}\left\{
    \sigma_{j} S_{ji} b_{ji}^{T}
  - \sigma_{i} S_{ij} b_{ij}^{T}
\right\}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$.

Eqns (\ref{eq:momFlux}) and (\ref{eq:bFlux}) are solved in advective form in which the mass transfer terms are more complicated: \remove[HW]{due to the mass transfer terms in \protect\ref{eq:sigma}}
\begin{eqnarray}
\frac{D_{i}\mathbf{u}_{i}}{Dt}+\nabla P_{i}
& = &
b_{i}\mathbf{k}+\nu\nabla^{2}\mathbf{u}_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(\mathbf{u}_{ji}^{T}-\mathbf{u}_{i}\right)-S_{ij}(\mathbf{u}_{ij}^{T}-\mathbf{u}_{i})-\mathbf{D}_{ij}\right\}
\label{eq:mom}\\
\frac{D_{i}b_{i}}{Dt}
& = &
\alpha\nabla^{2}b_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(b_{ji}^{T}-b_{i}\right)-S_{ij}\left(b_{ij}^{T}-b_{i}\right)\right\} \label{eq:b}
\end{eqnarray}
where $D_{i}\big/Dt=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$ is the total derivative with respect to fluid $i$,

Finally the multi-fluid version of incompressibility is
\begin{equation}
\sum_{i}\nabla\cdot\sigma_{i}\mathbf{u}_{i}  = \nabla\cdot\sum_{i}\sigma_{i}\mathbf{u}_{i}=0.
\label{eq:divFree}
\end{equation}
